Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-06-19 src/tex/ms.tex
Update ms.tex
Started on 2.2

% Put more focus on the joint optimization instead of having a set of more accurate 
% coefficients, since the method is the main point
In this paper, we investigate the possibility of further improving the accuracy
of IMRPhenomD by jointly optimizing all the fitting coefficients given NR
waveforms, and what constraints one may face when trying to further improve. We
find that simply by applying gradient descent algorithm, one can obtain a better
set of waveform coefficients, thus improving the accuracy of the model.
Furthermore, by comparing the accuracy of optimized and original waveforms, we
find that model-generated waveforms are very sensitive to their source
parameters. Specifically, IMRPhenomD favors certain parts of the parameter
space. This means IMRPhenomD introduces systematic bias to other GW analysis
tasks. This showcases the flaws of the ansatz and allows us to have a deeper
understanding of Phenom models.

The rest of the paper is structured as follows: In Sec.~\ref{sec:method}, we
review the parameterization of the IMRPhenomD model and the mismatch function
that is used as an objective function for the calibration, followed by outlining the
specific optimization scheme used for recalibration. In
Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
of optimized waveforms with original waveforms. We also show how the
optimization result differs with waveforms of different source parameters. In
Sec.~\ref{sec:discussion}, we address the difference between our calibrating
procedure with \citep{khan2016frequency}. We also explain how reduced spin
parameterization affects the accuracy of the model.

\section{Method} \label{sec:method}

@@

In order to recalibrate the model, we have to understand what parameters the model has.
Here we give a succinct summary of the IMRPhenomD model and the relevant parameters.
For interested readers, please refer to \citep{khan2016frequency} for more details on construction of the model.

The IMRPhenomD model is constructed by combining three individually fitted parts
into one coherent waveform model, which consists of the inspiral, intermediate, and merger-ringdown part, 
\begin{align}\label{eq:joint_waveform}
	h(f,\theta,\Lambda^i)=h_{\mathrm{ins}}(f,\theta,\Lambda^i) + h_{\mathrm{int}}(f,\theta,\Lambda^i) + h_{\mathrm{rd}}(f,\theta,\Lambda^i).
\end{align}
Instead of fitting the strain, which is a highly oscillatory function that is
difficult to fit, the amplitude and phase are fitted since they are smoother
functions. In each part, the amplitude and phase are made using simple functions
of frequency such as polynomials or lorentzians. Specifically, the merger-ringdown 
amplitude is fitted by a lorentzian and the other parts are fitted using polynomials. 
\begin{equation}\label{eq:amplitude}
\begin{aligned}
	A_0&\equiv\sqrt{\frac{2\eta}{3\pi^{1/3}}}f^{-7/6}\\
	A_{\mathrm{ins}}(f;\theta)&=A_{\mathrm{PN}}(f;\theta)+A_0\sum_{i=1}^3\rho_if^{(6+i)/3} \\
	A_{\mathrm{int}}(f;\theta)&=A_0(\delta_0+\delta_1f+\delta_2f^2+\delta_3f^3+\delta_4f^4) \\
	A_{\mathrm{rd}}(f;\theta)&=A_0\left[\gamma_1\frac{\gamma_3f_{\mathrm{damp}}}{(f-f_{\mathrm{RD}})^2+(\gamma_3f_{\mathrm{damp}})^2}e^{-\frac{\gamma_2(f-f_{\mathrm{RD}})}{\gamma_3f_{\mathrm{damp}}}}\right],  
\end{aligned}
\end{equation}
where $A_{\mathrm{PN}}$ is the post-newtonian expansion of the insprial amplitude up to order $A_0f^2$, $f_{\mathrm{damp}}$ is the damping frequency, and $f_{\mathrm{RD}}$ is the frequency at ringdown. The ansatzes for phase can be found in \citep{khan2016frequency}. These simple analytic functions
consists of parameters $\Lambda^i=\{\rho_i,\delta_i,\gamma_i\}$, which are defined as follows. 
\begin{align} \label{eq:Lambda}
	\Lambda^i&=\lambda_{00}^i+\lambda_{10}^i\eta \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)(\lambda_{01}^i+\lambda_{11}^i\eta+\lambda_{21}^i\eta^2) \nonumber \\ 
	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2),
\end{align}
where $\lambda$ are fitting coefficients obtained during calibration, $\eta$ is
the symmetric mass ratio, and $\chi_{\mathrm{PN}}$ is the post-Newtonian spin
parameter, which is defined as 
\begin{align}
	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2).
\end{align}
Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
respectively. Finally, the individual segments are first connected directly using 
step functions. Then, by fixing coefficients in the intermediate segment, one can make 
the final waveform is continuous in its first derivative.

%\kw{Need to work on this paragraph}
Combining Eq.~\ref{eq:joint_waveform},~\ref{eq:amplitude}~and~\ref{eq:Lambda}, 
we can see that the entire waveform is non-linear in $\lambda$. A slightly 
inaccurate set of $\lambda$ can significantly affect the shape of the generated 
waveforms. Thus, having a set of accurate waveform coefficients is important 
and fundamental to having accurate GW models. Generally, waveform coefficients 
are obtained by calibrating with NR waveforms, which are waveforms computed 
using NR simulations. In the case of \citep{khan2016frequency}, they first 
obtain a set of $\Lambda$ by fitting model generated waveforms to Eq.~\ref{eq:amplitude} 
and the phase ansatzes. Repeating with different NR waveforms, they obtain multiple 
sets of $\Lambda$, and $\lambda$ are subsequently found by fitting against 
Eq. \ref{eq:Lambda}. Since the fitting procedure is done in a piece-wise manner, 
the correlations between different segments are omitted, which could limit the 
accuracy of the model. Also, since fitting was performed before connecting 
individual segments, the final waveform does not guarantee to achieve the optimal 
waveform. The connecting procedure alters the previously fitted waveform. 
Hence, the model generated waveforms contains additional inaccuracies.

Instead, we recalibrate coefficients jointly, which we can remove inaccuracies
and biases discussed above, and can improve model accuracy. In the past, due to
the complex nature of GW strains and piece-wise formalism of IMRPhenomD,
non-linear fitting was difficult to be performed in optimizing coefficients.
Hence, piece-wise optimization was done to obtain coefficients. However, with
{\ripple} and AD from \jax, gradients of IMRPhenomD can be easily obtained, thus
allowing the use of gradient-based algorithms for us to recalibrate the model.

@@

In order to recalibrate the model, we need to define an objective function that
quantify the performance of the model. A common choice for such metric is the
\textit{mismatch} function \citep{husa2016frequency}. It is defined as
\begin{align} \label{eq:mismatch}
	\mathcal{M}(h_1, h_2)=1-\max_{t_0, \phi_0}\langle \hat{h}_1, \hat{h}_2\rangle,
\end{align}
where $h_{1,2}$ are the two GW waveform we are comparing, and $t_0$ and $\phi_0$
are time shift and phase shift respectively. The $\langle h_1, h_2 \rangle$ is
commonly referred as the inner product, which is defined as 
\begin{align}\label{eq:inner_prod}
	\langle h_1, h_2 \rangle = 4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
\end{align}
where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
$S_n(f)$ is the power spectral density(PSD) of noise from the instrument,
$f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are the relevant maximum and minimum
frequencies for the integration. $t_0$ and $\phi_0$ are maximized by performing
regression between $h_1$ and $h_2$. The mismatch is a quantity that is closely
related to the mean square error (MSE) between the two waveforms, as it is often
considered a weighted version of the MSE.

@@

\begin{figure}[t]
	\script{intrin_space.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
	\caption{Distribution of training and testing waveforms in $q-\chi_{PN}$
	space. The blue dots represent the waveforms used in training and the orange
	stars represent the waveforms used in testing.\kw{Change the legend, and include the extra points. Change the training point to be star and testing point to be dot.}}
	\label{fig:intrin_space}
\end{figure}

@@

To evaluate how well the optimization procedure generalize to waveforms that are
not in the training set, we evaluate the mismatch between the fine-tuned model
and the data provided in the SXS catalog for 536 NR waveforms. We select
waveforms that share the same part of the parameter space with the training
waveform, i.e. the waveforms with negligible eccentricity (${e<2\times10^{-3}}$)
and precession (${\chi_{x,y}<5\times10^{-3}}$). Fig.~\ref{fig:intrin_space}
shows how the training and testing waveforms are distributed in the
$q-\chi_{PN}$ space.

To illustrate the effect of optimiziation on an individual waveform level, we
compare the mismatch of a particular waveform before and after optimization with
respect to the NR waveform taken directly from the SXS catalog in Fig.
\ref{fig:0154}. Compared to the original IMRPhenomD waveform, the
optimized waveform has smaller residual from NR waveform both in amplitude and
phase, particularly in the inspiral region, where the amplitude displays a
$50\%$ reduction in error. For a fair comparison, we selected one of the testing
waveforms from the catalog presented in \citep{khan2016frequency}.

With the purpose of improving downstream tasks such as parameter estimation in
mind, the more relevant metric of improvement is the distribution of improvement
in mismatch over the entire training dataset. Fig. \ref{fig:q148} shows the
distribution of log-mismatch for the testing waveforms before and after the
optimization procedure. For generality, we use a constant PSD in our loss
function. One can see the distribution have been skewed toward lower mismatch,
where the peak of the distribution is shifted by approximately an order of
magnitude, and a median mismatch is reduced by 50\%. When using
$\mathcal{L}_{\mathrm{fl}}$, we observe a similar improvement with a 22.9\%
decrease in the median.

@@

Note that the IMRPhenomD model was initially constructed and fitted using the
{\zdethp} weighted mismatch. To ensure a fair comparison, we apply the same
methods using the {\zdethp} PSD exhibit superior improvement compared to the
unweighted mismatch. We find no significant qualitative and quantitative
difference between the two PSDs in the resulting distribution of mismatches.

To understand whether additional training data can further improve the
performance of the model, we include waveforms that are not presence in the
original IMRPhenomD calibration in our training dataset with parameters
tabulated in Tab. \ref{tab:q1248}. We specifically choose to use $q=2$ events
since we have abundant $q=2$ NR waveforms to validate the final result. The new
set of coefficients generated from this optimization process yields only
marginal improvements in the newly produced waveforms, as seen in Fig.
\ref{fig:q148_q1248_compare}. The high mismatch tail of the optimized
distribution remains comparable to the original distribution, indicating that
the original dataset is sufficient for this task. Similarly, utilizing the
{\zdethp} PSD to optimize the loss function with additional waveforms results in 
similar level of improvement.

To investigate the performance of recalibration over the source parameter space,
we plot the improvement of mismatch in log projected over the parameter space of
$q$ vs. $\chi_{\mathrm{PN}}$ in Fig. \ref{fig:ps_q148_qchi}. On the mass ratio
axis, we can see the waveforms with $q\lq4$ show most consistent average
improvement. This is because that part of the parameter space is better covered
by the training waveform, as compared to testing waveforms from $q=4$ to $q=8$,
where some testing waveforms lie outside the source parameter space covered by
the training set of waveforms. On the spin axis, we can see the waveform with
$\chi_{\textrm{PN}}$ close to zero show the most consistent improvement, and we
start deviate from 0, there could be fluctuations in the average improvement. On
top of that, in the $q\lq4$ region, there seems to be more consistent
improvement of waveforms with $\chi_{\textrm{PN}}<0$. This can be understood as
since the original waveform was developed for aligned spin system, the waveforms
with $\chi_{\textrm{PN}}$ is less well-fitted \kw{Varify this.}, hence there
could be more room for improvement. To show this, we plot the parameter space of
$\chi_1$ vs. $\chi_2$ in Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal
axis, i.e. $\chi_1\approx\chi_2$, show good mismatch improvements as discussed
above. Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right
($\chi_1>0>\chi_2$) regions respond to optimization differently. In the top-left
region, waveforms generally improve slightly with along optimization. However,
waveforms in the bottom-right region do not improve after optimization. Some
waveforms even turned worse after optimization.

 Given that the waveform model's ansatz may not be entirely compatible with NR,
and the optimization procedure is carried out over a distribution of waveforms
with varying source parameters, it is conceivable that different parts of the
source parameter space may not share the same set of optimal IMRPhenomD
parameters, mean there could be some trade-offs in accuracy between different
parts of the parameter space. If this is the cause of the lack of improvement in
the high mismatch tail of the distribution, segmenting the parameter space into
smaller subspaces should alleviate this problem. On the other hand, if the
ansatz lacks the correct parameterized form to capture the NR waveforms'
behavior as a function of the source parameters, the results will always be
biased, and we should not expect any improvement, even if we segment the
parameter space during training.

@@

We divided the parameter space into four regions to analyze the effect of the
recalibration procedure on each region separately
(Fig.~\ref{fig:ps_q148_quadrant}). The training waveforms used for fitting in
this scenario are listed in Tab.~\ref{tab:quadrants}. The top-left and
bottom-right regions have limited data for $q>4$, hence the results are only
valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that
equal-spin waveforms lying on the diagonal has great improvement. The
improvement above the diagonal is also substantial compared to recalibrate using
all waveforms at once, barring a few caused by some testing waveforms having
$q>4$. This can be seen clearer in Fig.~\ref{fig:all_quadrants}, which we see
that the optimized histogram shifts downward uniformly. On the other hand, the
improvement below the diagonal is negligible, indicating our method struggles to
further improve on top of the original IMRPhenomD waveform. Interestingly, we
can see in the top-right and bottom-left quadrant, the lower right corner shows
systematic worsening similar to the case of all waveforms. This hinted in
general, the IMRPhenomD waveform ansatz does not fit well with waveforms with
anti-aligned spin where the primary spin is positive.

\begin{figure}[t]
	\script{all_quadrants.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
	\caption{Distributions of mismatches in the top-left (Top) and bottom-right
	(Bottom) regions. We use a constant noise spectrum to calculate mismatch and
	$\mathcal{L}_{\mathrm{mean}}$ as the loss function. Waveforms in the
	top-left region generally improves while waveforms in the bottom-right
	region has very little improvement, as indicated by the median (dashed
	lines). \kw{Is the original referring to original IMRPhenomD or optimized
	using everything? Can we have a reference for optimizing using everything?}}
	\label{fig:all_quadrants}
\end{figure}

@@

Furthermore, we investigated how the source parameters affect the
improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a
certain preference for waveforms lying in the top-left region while the bottom-right 
region hardly improved. To further test this result, we recalibrate
waveforms in separate regions in parameter space. From Fig.~\ref{fig:ps_q148_quadrant}, 
we can see that this recalibration process gives further
improvement to the top-left region while the bottom-right region only have little 
improvement. This indicates that the ansatz has limited match in this region and 
does not fit most waveforms in this region, hence introduces bias to downstream GW 
analysis. This phenomenon is mainly due to the reduced spin approximation used in 
parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.

>>>>

% Put more focus on the joint optimization instead of having a set of more accurate 
% coefficients, since the method is the main point
In this paper, we investigate the possibility of further improving the accuracy
of IMRPhenomD by jointly optimizing all the fitting coefficients given NR
waveforms, and what constraints one may face when trying to further improve. We
find that simply by applying gradient descent algorithm, one can obtain a better
set of waveform coefficients, thus improving the accuracy of the model.
Furthermore, by comparing the accuracy of optimized and original waveforms, we
find that model-generated waveforms are very sensitive to their intrinsic
parameters. Specifically, IMRPhenomD favors certain parts of the parameter
space. This means IMRPhenomD introduces systematic bias to other GW analysis
tasks. This showcases the flaws of the ansatz and allows us to have a deeper
understanding of Phenom models.

The rest of the paper is structured as follows: In Sec.~\ref{sec:method}, we
review the parameterization of the IMRPhenomD model and the mismatch function
that is used as an objective function for the calibration, followed by outlining the
specific optimization scheme used for recalibration. In
Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches
of optimized waveforms with original waveforms. We also show how the
optimization result differs with waveforms of different intrinsic parameters. In
Sec.~\ref{sec:discussion}, we address the difference between our calibrating
procedure with \citep{khan2016frequency}. We also explain how reduced spin
parameterization affects the accuracy of the model.

\section{Optimization Method} \label{sec:method}

In this section, we first briefly review the construction of the IMRPhenomD model and discuss
how the calibration parameters enter the waveform.
We then outline the mismatch and how it can be used as a loss fucntion. 
Finally we discuss the gradient descent algorithm and our stopping criterion.

@@

% Paragraphs we need:
% 1. What is the model
% 2. How do the parameters enter the model
We start by giving a succinct summary of the IMRPhenomD model and the relevant parameters.
Interested readers should refer to \citep{khan2016frequency} for more details.

Aligned-spin, frequency-domain waveform models (such as IMRPhenomD) can be written as a
combination of amplitude and phase functions ($A$ and $\phi$ respectively):
\begin{align}\label{eq:}
	h(f,\theta,\Lambda) = A(f,\theta,\Lambda)e^{-i\phi(f,\theta,\Lambda)}\,,
\end{align}
where $f$ is the frequnecy, $\theta$ are the intrinsic parameters of the binary, and $\Lambda$ is a set of additional
parameters which will be discussed below. 
The phase and amplitude functions are then split into three sections which represent the
inspiral, intermediate, and merger-ringdown (MR) parts of the waveform. 
During inspiral, $A$ and $\phi$ are known analytically from post-Newtonian (PN) theory;
IMRPhenomD uses the TaylorF2 model \citep{Buonanno:2009zt, Arun:2004hn} which is known up to 3.5PN order.
To model the intermediate and MR regions, IMRPhenomD (and all waveforms in the IMRPhenom family)
uses a series of parameterizations\footnote{
	The parameterizations for both the ampltude and phase functions can be found in \citep{khan2016frequency}.
} 
which depend purely on $\Lambda$ and can be calibrated to numerical relativity (NR) simulations.
The three sections are then \textit{stictched} together using step functions.
Importantly, the parameterizations are chosen such that they can be made $\mathcal{C}^1$ continuous at the
boundary between each section.

In practice the $\Lambda$ parameters are fit for each section independently i.e., intermediate coefficients are fit whilst ignoring the MR region.
Finally, to map the grid of tuned $\Lambda$ parameters back to the intrinsic parameter space, IMRhenomD uses the polynomial function:
\begin{align} \label{eq:Lambda}
	\Lambda^i&=\lambda_{00}^i+\lambda_{10}^i\eta \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)(\lambda_{01}^i+\lambda_{11}^i\eta+\lambda_{21}^i\eta^2) \nonumber \\ 
	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2)\,,
\end{align}
where the $\lambda$'s are the fitting coefficients we are going to optimize below, $\eta$ is
the symmetric mass ratio, and $\chi_{\mathrm{PN}}$ is the post-Newtonian spin
parameter, which is defined as 
\begin{align}
	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2)\,.
\end{align}
Here, $m_{1,2}$ and $\chi_{1,2}$ are the primary and secondary mass and spin,
respectively.

% More precisely, $\Lambda$ by fitting model generated waveforms to Eq.~\ref{eq:amplitude} 
% and the phase ansatzes. Repeating with different NR waveforms, they obtain multiple 
% sets of $\Lambda$, and $\lambda$ are subsequently found by fitting against 
% Eq. \ref{eq:Lambda}. Since the fitting procedure is done in a piece-wise manner, 
% the correlations between different segments are omitted, which could limit the 
% accuracy of the model. Also, since fitting was performed before connecting 
% individual segments, the final waveform does not guarantee to achieve the optimal 
% waveform.
% Finally, the individual segments are first connected directly using 
% step functions. Then, by fixing coefficients in the intermediate segment, one can make 
% the final waveform is continuous in its first derivative.

% In order to recalibrate the model, we have to understand what parameters the model has.
% Here we give a succinct summary of the IMRPhenomD model and the relevant parameters.
% For interested readers, please refer to \citep{khan2016frequency} for more details on construction of the model.

% The IMRPhenomD model is constructed by combining three individually fitted parts
% into one coherent waveform model, which consists of the inspiral, intermediate, and merger-ringdown part, 
% \begin{align}\label{eq:joint_waveform}
% 	h(f,\theta,\Lambda^i)=h_{\mathrm{ins}}(f,\theta,\Lambda^i) + h_{\mathrm{int}}(f,\theta,\Lambda^i) + h_{\mathrm{rd}}(f,\theta,\Lambda^i).
% \end{align}
% Instead of fitting the strain, which is a highly oscillatory function that is
% difficult to fit, the amplitude and phase are fitted since they are smoother
% functions. In each part, the amplitude and phase are made using simple functions
% of frequency such as polynomials or lorentzians. Specifically, the merger-ringdown 
% amplitude is fitted by a lorentzian and the other parts are fitted using polynomials. 
% \begin{equation}\label{eq:amplitude}
% \begin{aligned}
% 	A_0&\equiv\sqrt{\frac{2\eta}{3\pi^{1/3}}}f^{-7/6}\\
% 	A_{\mathrm{ins}}(f;\theta)&=A_{\mathrm{PN}}(f;\theta)+A_0\sum_{i=1}^3\rho_if^{(6+i)/3} \\
% 	A_{\mathrm{int}}(f;\theta)&=A_0(\delta_0+\delta_1f+\delta_2f^2+\delta_3f^3+\delta_4f^4) \\
% 	A_{\mathrm{rd}}(f;\theta)&=A_0\left[\gamma_1\frac{\gamma_3f_{\mathrm{damp}}}{(f-f_{\mathrm{RD}})^2+(\gamma_3f_{\mathrm{damp}})^2}e^{-\frac{\gamma_2(f-f_{\mathrm{RD}})}{\gamma_3f_{\mathrm{damp}}}}\right],  
% \end{aligned}
% \end{equation}
% where $A_{\mathrm{PN}}$ is the post-newtonian expansion of the insprial amplitude up to order $A_0f^2$, $f_{\mathrm{damp}}$ is the damping frequency, and $f_{\mathrm{RD}}$ is the frequency at ringdown.

%\kw{Need to work on this paragraph}
Although initially independent, the stiching procedure means that each section
of the waveform intrinsically depends on the full set of $\lambda$'s. 
A slightly inaccurate set of $\lambda$'s can therefore lead to inaccuracies in
the generated waveforms. 
Thus, the calibration of these coefficients is crucial to the accuracy
of IMRPhenom GW models. 
Importantly, since the fitting was performed on the individual segments,
the final waveform is not guarenteed to have $\lambda$'s close to global minima.

% Generally, waveform coefficients 
% are obtained by calibrating with NR waveforms, which are waveforms computed 
% using NR simulations. In the case of \citep{khan2016frequency}, they first 
% obtain a set of $\Lambda$ by fitting model generated waveforms to Eq.~\ref{eq:amplitude} 
% and the phase ansatzes. Repeating with different NR waveforms, they obtain multiple 
% sets of $\Lambda$, and $\lambda$ are subsequently found by fitting against 
% Eq. \ref{eq:Lambda}. Since the fitting procedure is done in a piece-wise manner, 
% the correlations between different segments are omitted, which could limit the 
% accuracy of the model. Also, since fitting was performed before connecting 
% individual segments, the final waveform does not guarantee to achieve the optimal 
% waveform. The connecting procedure alters the previously fitted waveform. 
% Hence, the model generated waveforms contains additional inaccuracies.

At the time of construction this piece-wise approach was necessary since
$\lambda$ has 209 components, making the fitting to NR simulations computationally prohibitive.
Here, for the first time we recalibrate the $\lambda$ coefficients jointly. 
This is made possible by the use of gradient-based optimization algorithms,
enabled by AD from \jax\, and {\ripple}, which are significatnly more efficient in high dimensions.
% , gradients of IMRPhenomD can be easily obtained, thus
% allowing the use of gradient-based algorithms for us to recalibrate the model.

@@

In order to optimize the coefficients, we need to define a loss function that
quantifies the difference between waveform model and the target NR simulations which we want to match.
Here we adopt a quantity commonly used in GW physics called the \textit{mismatch} function \citep{husa2016frequency}. 
It is defined as
\begin{align} \label{eq:mismatch}
	\mathcal{M}(h_1, h_2)=1-\max_{t_0, \phi_0}\langle \hat{h}_1, \hat{h}_2\rangle,
\end{align}
where $h_{1,2}$ are the two GW waveforms we are comparing, and $t_0$ and $\phi_0$
are a relative time and phase shift respectively. 
The inner product, $\langle h_1, h_2 \rangle$, is defined as 
\begin{align}\label{eq:inner_prod}
	\langle h_1, h_2 \rangle = 4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
\end{align}
where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain,
$S_n(f)$ is the power spectral density (PSD), $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are
the relevant maximum and minimum frequencies for the integration.
We note here that the mismatch can be viewed as the mean square error (MSE) between the
two waveforms.

\te{Up to here}

@@

\begin{figure}[t]
	\script{intrin_space.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
	\caption{Parameter space with mass ratio $q$ against post-newtonian spin parameter $\chi_{\mathrm{PN}}$. Orange: Training waveforms; Blue: Testing waveforms}
	\label{fig:intrin_space}
\end{figure}

@@

To evaluate the effectiveness of the optimization, we analyze 536 NR waveforms 
from the SXS catalog in this study. We specifically select waveforms with 
negligible eccentricity (${e<2\times10^{-3}}$) and precession 
(${\chi_{x,y}<5\times10^{-3}}$) that are consistent with the constraints of 
the waveform model. As shown in Fig.~\ref{fig:intrin_space}, the intrinsic 
parameters of the chosen test waveforms fall within the parameter space defined 
by the training waveforms. This suggests that these test waveforms can be used 
for direct comparison with the original model.

To illustrate the improvement on an individual waveform level, we compare a
particular waveform before and after optimization in Fig. \ref{fig:0154}.
Compared to the original IMRPhenomD waveform, we can see the optimized waveform
has smaller residual from NR waveform both in amplitude and phase, particularly
in the inspiral region, where the amplitude displays a $50\%$ reduction in
error. For a fair comparison, we selected one of the testing waveforms from the
catalog presented in \citep{khan2016frequency}.

To quantify the effect of optimization over the entire dataset, We evaluate the
mismatch of all testing waveforms using a constant PSD weighted loss function,
$\mathcal{L}_{\mathrm{mean}}$. We present the resulting distribution in Fig.
\ref{fig:q148}. The peak of the distribution has shifted towards a lower
mismatch, with a decrease of almost one order of magnitude and a 50\% reduction
in the median. When using $\mathcal{L}_{\mathrm{fl}}$, we observe a similar
improvement with a 22.9\% decrease in the median. Even though the distribution
lacks a clear peak due to the problem discussed in Section
\ref{subsec:optimization}, both distributions show observable improvement.

@@

%\kw{Which plot is this paragraph referring to?}
Applying the same methods, we will see that the distributions of mismatches
calculated using the {\zdethp} PSD exhibit superior improvement compared to the 
unweighted mismatch.
The shape of the distribution is similar to that shown in Fig. \ref{fig:q148}.
This result is expected, as the IMRPhenomD model was initially constructed and
fitted using the {\zdethp} weighted mismatch. Consequently, the model is
anticipated to closely fit the NR waveforms while incorporating the influence of
the {\zdethp} PSD, rather than a constant PSD.

Motivated by the successful improvement of the waveforms, we expand the training
dataset to optimize additional waveforms listed in Tab. \ref{tab:q1248}. We 
specifically choose to use $q=2$ events since we have abundant $q=2$ NR waveforms 
to validate the final result. The new set of coefficients generated from this
optimization process yields only marginal improvements in the newly produced
waveforms, as seen in Fig. \ref{fig:q148_q1248_compare}. The high mismatch tail
of the optimized distribution remains comparable in length and endpoint to the
original distribution, indicating that our procedure is incapable of improving
these waveforms. Similarly, utilizing the {\zdethp} PSD to optimize the loss
function with additional waveforms leads to similar improvement in the resulting
distribution.

% We believe that either the ansatz does not suit waveforms within certain
% regions of parameter space or the set of optimized coefficients falls to a
% minimum where they do not fit some testing waveforms. If in some region of
% parameter space, the ansatz does not fit well with NR waveforms, it would
% never be able to have consistent improvement under optimization. Instead, it
% would surf around the loss manifold with random fluctuations in mismatches. If
% the set of coefficients does not fit, then dividing the parameter space into
% regions and fitting separate sets of coefficients should give better results.
Although most of the waveforms show improvement in figure \ref{fig:q148}, the
high mismatch tail of the distribution remains unaffected. Given that the
waveform model's ansatz may not be entirely compatible with NR, and the
optimization procedure is carried out over a distribution of waveforms with
varying intrinsic parameters, it is conceivable that some trade-offs in accuracy
exist between different parts of the parameter space. If this is the cause of
the lack of improvement in the high mismatch tail of the distribution,
segmenting the parameter space into smaller subspaces should alleviate this
problem. On the other hand, if the ansatz lacks the correct parameterized form
to capture the NR waveforms' behavior as a function of the intrinsic parameters,
the results will always be biased, and we should not expect any improvement,
even if we segment the parameter space during training.

Since we know intrinsic parameters play an important role in the ansatz, we
would like to investigate how intrinsic parameters affect the recalibration
process. First, we plot the parameter space of $q$ vs. $\chi_{\mathrm{PN}}$ in
Fig. \ref{fig:ps_q148_qchi}. We can see near non-spinning waveforms demonstrate
more consistent improvement, since the ansatz are developed base on non-spinning 
waveforms. Also, the original coefficients were fitted using NR waveforms with
equal or similar spin, hence the model prefers waveforms with similar spin. On 
the other hand, spinning waveforms can either improve or worsen in terms of 
mismatch base on their intrinsic parameters. To further discuss the behavior of 
non-spinning waveforms, we plot the parameter space of $\chi_1$ vs. $\chi_2$ in
Fig.~\ref{fig:ps_q148}. Waveforms along the diagonal axis, i.e. 
$\chi_1\approx\chi_2$, show good mismatch improvements as discussed above. 
Meanwhile, the top-left ($\chi_1<0<\chi_2$) and bottom-right ($\chi_1>0>\chi_2$) 
regions respond to optimization differently. In the top-left 
region, waveforms generally improve slightly with along optimization. However, waveforms 
in the bottom-right region do not improve after optimization. Some waveforms even 
turned worse after optimization.

@@

We divided the parameter space into four regions to analyze the effect of the
recalibration procedure on each region separately (Fig.~\ref{fig:ps_q148_quadrant}). 
The training waveforms used for fitting are listed in Tab.~\ref{tab:quadrants}. The
top-left and bottom-right regions have limited data for $q>4$, hence the results are
only valid up to $q\lesssim4$. From Fig.~\ref{fig:ps_q148_quadrant}, we observe that 
equal-spin waveforms lying on the diagonal has great improvement. Above the diagonal, 
the improvement is significant, except for a few defects caused by some testing waveforms 
having $q>4$. This can be seen clearer in Fig.~\ref{fig:all_quadrants}, which we see 
that the optimized histogram shifts downward uniformly. Below the diagonal, waveforms 
have negligible improvement (Fig.~\ref{fig:all_quadrants}~and~\ref{fig:ps_q148_quadrant}), 
indicating the original set of coefficients is the optimal set of coefficients, and 
cannot be further improved. Hence, we can deduce IMRPhenomD has better match with the 
top-left region and does not fit well with waveforms lying in the bottom-right 
region. Thus, the ansatz restricts additional improvements and further optimizing in a 
smaller region does not generally improve the results.

\begin{figure}[t]
	\script{all_quadrants.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/all_quadrants.pdf}
	\caption{Distributions of mismatches in the top-left (Top) and bottom-right (Bottom) regions.
	We use a constant noise spectrum to calculate mismatch and
	$\mathcal{L}_{\mathrm{mean}}$ as the loss function. Waveforms in the top-left region generally improves while waveforms in the bottom-right region has very little improvement, as indicated by the median (dashed lines).}
	\label{fig:all_quadrants}
\end{figure}

@@

Furthermore, we investigated how the intrinsic parameters affect the
improvement. Fig.~\ref{fig:ps_q148} shows that the optimization procedure has a
certain preference for waveforms lying in the top-left region while the bottom-right 
region hardly improved. To further test this result, we recalibrate
waveforms in separate regions in parameter space. From Fig.~\ref{fig:ps_q148_quadrant}, 
we can see that this recalibration process gives further
improvement to the top-left region while the bottom-right region only have little 
improvement. This indicates that the ansatz has limited match in this region and 
does not fit most waveforms in this region, hence introduces bias to downstream GW 
analysis. This phenomenon is mainly due to the reduced spin approximation used in 
parameterizing the ansatz, where degeneracies between $\chi_1$ and $\chi_2$ are introduced.
